03_wrangle

### 2020-08-28 analysis/03_datawrangling.Rmd
working on better tables

parentalcaregenes <- GOgenesLong %>% 
  pivot_wider(
    names_from = GO,
    values_from = gene,
    values_fill = list(gene = "")) %>%
  left_join(geneids, by = c("geneid","NCBI")) %>%
  dplyr::rename("GO"= "parentalbehavior" )
head(parentalcaregenes)

@@

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  #pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               #names_to = "samples", values_to = "counts") 
  pivot_longer(cols = L.G118_female_hypothalamus_control.NYNO:y98.g54_female_hypothalamus_m.hatch, 
               names_to = "samples", values_to = "counts") 
allvsd %>% select(-file_name) %>% head()

tail((allvsd))
```

@@

wideDEGs <- tempDEGs %>%
  mutate(lfc = round(lfc, 2),
         padj =  round(padj,2)) %>%
  mutate(sextissue = paste(sex, tissue, sep = "_"),
         lfcpadj = paste(lfc, padj, sep = ", "))  %>%
  dplyr::select(sextissue,comparison, gene, lfcpadj) %>%
  pivot_wider(names_from = comparison, values_from = lfcpadj)

wideDEGscandidates <- wideDEGs %>%
  filter(gene %in% GOgenesLong$gene)
wideDEGscandidates

wideDEGsBldg <- wideDEGscandidates %>%
  select(sextissue,gene, contains("bldg"))
wideDEGsBldg

wideDEGsControl <- wideDEGscandidates %>%
  select(sextissue,gene, contains("control"))
wideDEGsControl

# sequential in 2 or more comparisons
wideDEGsSequential <- wideDEGscandidates %>%
  select(sextissue,gene, "control_bldg", "bldg_lay" #, 
         #"lay_inc.d3", "inc.d3_inc.d9",
         #"inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9"
         ) %>%
  mutate(totalNA = rowSums(is.na(.))) %>%
  filter(totalNA < 8) %>%
  arrange(totalNA, gene, sextissue) %>%
  arrange( gene, sextissue) %>%
  select(-totalNA) %>% 
  replace(., is.na(.), "NS")
wideDEGsSequential

@@

removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
early <- c("early_inc.d9", "hatch_early")
late <- c("hatch_prolong", "inc.d17_prolong")

removalDEGs <- makemanipulationDEGtables(removalcomps)
hatchDEGs <- makemanipulationDEGtables(early)
prolongDEGs <- makemanipulationDEGtables(late)

@@

```{r write}
#write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
#write.csv(candidatevsd, "../results/03_candidatevsd.csv")
#write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)

#write.csv(removalDEGs, "../results/removalDEGs.csv")
#write.csv(hatchDEGs, "../results/hatchDEGs.csv")
#write.csv(prolongDEGs, "../results/prolongDEGs.csv")

#write.csv(manipDEGs, "../results/manipDEGs.csv")

>>>>

parentalcaregenes <- GOgenesLong %>% 
  pivot_wider(
    names_from = GO,
    values_from = gene,
    values_fill = list(gene = "")) %>%
  left_join(geneids, by = c("geneid","NCBI")) %>%
  dplyr::rename("GO"= "parentalbehavior" )
head(parentalcaregenes)
```

@@

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  #pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               #names_to = "samples", values_to = "counts") 
  pivot_longer(cols = L.Blu13_male_hypothalamus_control.NYNO:y98.o50.x_male_hypothalamus_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd %>% select(-file_name) %>% head()

tail(names(allvsd))
```

@@

# sequential in 2 or more comparisons
wideDEGscandidates <- tempDEGs %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G"),
         group = paste(sex, tissue, sep = "")) %>%
  mutate(res = paste("(", posneg, ")", sep = "")) %>%
  mutate(compres = paste(group, res, sep = "")) %>%
  group_by(gene, comparison) %>%
  summarize(results = str_c(compres, collapse = "; ")) %>%
  pivot_wider(names_from = comparison, values_from = results) 
wideDEGscandidates

unique(wideDEGscandidates$compres)
```

@@

removalcomps <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2")
earlys <- c("inc.d9_early", "inc.d17_early", "hatch_early", "n5_early")
prolongs <- c("inc.d9_prolong","inc.d17_prolong", "hatch_prolong",  "n5_prolong")
extends <- c("inc.d9_extend","inc.d17_extend", "hatch_extend",  "n5_extend")

removalDEGs <- makemanipulationDEGtables(removalcomps)
earlyDEGs <- makemanipulationDEGtables(earlys)
prolongDEGs <- makemanipulationDEGtables(prolongs)
extendDEGs <- makemanipulationDEGtables(extends)

@@

```{r write}
write.csv(allDEG, "../results/03_allDEG.csv", row.names = F)
write.csv(candidatevsd, "../results/03_candidatevsd.csv")
write.csv(wideDEGsSequential, "../results/table1.csv", row.names = F)
write.csv(manipDEGs, "../results/manipDEGs.csv")
```
